### 2011-11-04 doc/pldi2012/highlevel.tex
Removing the word conflict from the high level runtime description

\subsection{Dependence Analysis} 
\label{subsec:depanalysis}
The first step in the execution of a task is dependence analysis.  When
a task is registered with the high level runtime the task call includes
information about both the logical regions that the task will require
when it is executed as well as the read-write access and coherence properties
for each of the regions.  The runtime uses this information to perform
conflict detection with each of the previous tasks that have been registered
in the same parent task.  Conflict detection only needs to be performed at
the scope of an enclosing task because the semantics of the Legion programming
model ensure that all tasks can only use regions which are subregions of
regions that the parent task uses.  This implies that if a child task were
to have a conflict with task at a higher level of the task tree, then the
parent task would also have had a conflict.

Lemma: If a task {\tt t} with ancestor task {\tt p} conflicts with task {\tt t'}
a sibling task of {\tt p}, then tasks {\tt p} and {\tt t'} must conflict.

@@

When performing dependence analysis the runtime relies on its knowledge about
the relationships between logical regions and partitions.  We first describe
the the components to dependence analysis in sections \ref{subsec:regiontree},
\ref{subsec:cohconf}, and \ref{absinsts}.  We then describe the algorithm for performing 
dependence analysis on this data structures in section \ref{subsec:depdetect}.

\subsection{Region Trees}
\label{subsec:regiontree}
To detect dependences between the tasks, the runtime leverages its knowledge
about logical regions and their partitions.  To create a new logical region
or partition, the application must invoke the corresponding call in the
runtime.  As these calls occur, the runtime constructs a data strcture
called a {\em region tree}.  A region tree describes the relationship between
logical regions and partitions.  A region tree contains two types of nodes:
region nodes and partition nodes.  Every region tree is rooted with a region
node and alternates between region and partition nodes at each corresponding level.
A region node represents a specific logical region and tracks the set of
partitions of that logical region.  Similarly a partition node represents 
a specific partitioning of a logical region and keeps track of the logical regions 
which are subregions of the partition.

\subsection{Coherence Conflicts}
\label{subsec:cohconf}
To detect conflicts between any two tasks using a logical region, we leverage
the information contained in the access and cohrence properties provided by the programmer
for how the region is to be used.  Conflict checks are performed pair-wise between regions
to determine whether two tasks conflict with each other.  When checking for
conflicts we first examine the access property.  If both tasks are reads, then 
there is never a conflict.  If at least one of them is a write we use table
\ref{tab:conflicts} to determine whether there is a conflict.

\begin{center}
\begin{table}
\begin{tabular}{|c|c|c|c|c|} \hline
             & Exclusive & Atomic   & Simultaneous & Relaxed \\ \hline 
Exclusive    & Conflict  & Conflict & Conflict     & Conflict \\ \hline
Atomic       & Conflict  & Same Instance & Conflict & Conflict \\ \hline
Simultaneous & Conflict  & Conflict & Same Instance & None \\ \hline
Relaxed      & Conflict  & Conflict & None & None \\ \hline
\end{tabular}
\caption{Conflict detection cases}
\end{table}
\end{center}

\subsection{Abstract Instances}
\label{subsec:absinsts}

\subsection{Dependence Detection}
\label{subsec:depdetect}
When a task is registered, the runtime
will perform conflict detection for every region that the task requests with
prior tasks that have been registered in the same parent task context.  Referring
back to our analogy of the high level runtime as an out of order processor, this
corresponds to the wake-up logic that detects when an instruction is ready to be
executed.  Continuing this analogy, the regions that a task requires are annalogous
to the registers listed in an instruction.  However, this poses a difficulty for the
high level runtime as two logical regions are not necessarily independent.  In
order to do the conflict detection we rely on the region tree data structure to be
able to detect conflicts between tasks which require different logical regions.

@@

Once we have found the root region for the requested region, we then proceed to walk
down the region tree from the node corresponding to the root region to the node
for the requested region.  At each node, we determine if any potential conflicts exist.

For region nodes we first check to see if any tasks currently using this region
conflict based on the coherence proprties the two task require.  We describe coherence
conflicts in more detail in section \ref{subsec:cohconf}.  If a conflict is discovered
we record that the task must wait on the previous task before executing.  We also check
to see if any of the partitions of the region have been marked open.  If they have
we must {\em close} the partitions to detect any possible conflicts

\subsection{Correctness}

@@

\subsubsection{Instance Management}

>>>>

\subsection{Dependence Analysis} 
\label{subsec:depanalysis}
The first step in the execution of a task is dependence analysis.  When
a task is registered with the high level runtime the task call includes
information about both the logical regions that the task will require
when it is executed as well as the read-write access and coherence properties
for each of the regions.  The runtime uses this information to perform
dependence detection with each of the previous tasks that have been registered
in the same parent task.  Dependence detection only needs to be performed at
the scope of an enclosing task because the semantics of the Legion programming
model ensure that all tasks can only use regions which are subregions of
regions that the parent task uses.  This implies that if a child task were
to depend on a task at a higher level of the task tree, then the
parent task would also have had a dependence.

Lemma: If a task {\tt t} with ancestor task {\tt p} depends on task {\tt t'}
a sibling task of {\tt p}, then task {\tt p} must depend on task {\tt t'}.

@@

When performing dependence analysis the runtime relies on its knowledge about
the relationships between logical regions and partitions.  We first describe
the the components to dependence analysis in sections \ref{subsec:regiontree},
\ref{subsec:cohdep}, and \ref{absinsts}.  We then describe the algorithm for performing 
dependence analysis on this data structures in section \ref{subsec:depdetect}.

\subsubsection{Region Trees}
\label{subsec:regiontree}
To detect dependences between the tasks, the runtime leverages its knowledge
about logical regions and their partitions.  To create a new logical region
or partition, the application must invoke the corresponding call in the
runtime.  As these calls occur, the runtime constructs a data strcture
called a {\em region tree}.  A region tree describes the relationship between
logical regions and partitions.  A region tree contains two types of nodes:
region nodes and partition nodes.  Every region tree is rooted with a region
node and alternates between region and partition nodes at each corresponding level.
A region node represents a specific logical region and tracks the set of
partitions of that logical region.  Similarly a partition node represents 
a specific partitioning of a logical region and keeps track of the logical regions 
which are subregions of the partition.

\subsubsection{Dependencies From Coherence Properties}
\label{subsec:cohdep}
To detect dependencies between any two tasks using a logical region, we leverage
the information contained in the access and cohrence properties provided by the programmer
for how the region is to be used.  Dependency checks are performed pair-wise between regions
to determine whether one task depends on the other.  When checking for
dependencies we first examine the access property.  If both tasks are reads, then 
there is never a dependence.  If at least one of them is a write we use table
\ref{tab:depsdetect} to determine whether there is a dependence.

\begin{center}
\begin{table}
\begin{tabular}{|c|c|c|c|c|} \hline
             & Exclusive & Atomic   & Simultaneous & Relaxed \\ \hline 
Exclusive    & Dependence & Dependence & Dependence & Dependence \\ \hline
Atomic       & Dependence & Same Instance & Contingent & Contingent \\ \hline
Simultaneous & Dependence & Contingent & Same Instance & None \\ \hline
Relaxed      & Dependence & Contingent & None & None \\ \hline
\end{tabular}
\caption{Dependence detection cases when at least one access is a write.\label{tab:depsdetect}}
\end{table}
\end{center}
In this table, {\em Dependence} indicates that there will always be a dependence between
regions while {\em None} indicates that there will never be a dependence regardless
of mapping decisions.  In the case of {\em Same Instance} there is a dependence contingent
upon a mapping decision.  If both tasks map to the same physical instance (described in 
more detail in section \ref{subsec:instmang}) then there is no dependence, which if
they map to different instances there is a dependence as the system will have to generate
a copy between the two tasks running.

For the case of {\tt Contingent} there are two special cases that we handle.
One special case is Write-After-Read (WAR) dependencies.  For the WAR case, if neither coherence property
is {\em Exclusive} then we can ensure that different physical instances
are used to avoid an anti-dependence and increase parallelism.  If either property is {\em Exclusive} then the semantics
of the programming model mandate that we detect a dependence and serialize the task.

Another special case occurs if there is an {\tt Atomic} writer and either {\tt Simultaneous}
or {\tt Relaxed} reader.  In this case it is acceptable for the tasks to run in
parallel as it is valid for the reader to see updates form the writer, and the
reader will not violate the atomic access semantics of the writer.

\subsubsection{Abstract Instances}
\label{subsec:absinsts}
For the cases where tasks do not depend, they are allowed to create different physical
instances of the same logical region.  The decision to do this however does not
happen until the mapping phase of a task execution.  However, we need to have a place
holder that is capable of representing all the possible valid physical instances
from which to choose even if those instances do not exist because the prior
tasks have not been mapped.  We call these place holders {\tt abstract instances}.

Abstract instances represent the set of possible valid physical instances which a
task will be capable of selecting from or adding to when it is mapped.  The value
of abstract instances is that they allow us to remember which physical instances
are valuable while continuing to perform dependence detection on newly registered
tasks.  In the analogy to out-of-order processors abstract instances allow us
to perform the same operation as register-renaming of instructions to avoid
anti-dependences and to remember actual dependences even though our dependence detection
analysis may run ahead of the actual execution of tasks.

\subsubsection{Dependence Detection}
\label{subsec:depdetect}
When a task is registered, the runtime
will perform dependence detection for every region that the task requests with
prior tasks that have been registered in the same parent task context.  Since
tasks are registered in program order we will always perform dependence detection
in program order which guarantees correctness even if tasks are executed out-of-order.  
Performing dependence detection in order is also annalogous to how an out of order
processor works.  Out-of-order processors load instructions in order and only after
analyzing their instructions and register dependencies do they execute instructions
out of order.  The regions that a task requires are annalogous
to the registers listed in an instruction and the goal of the high level
runtime is to detect the same dependences and use them to enforce scheudling constraints
upon a task.  However, there is a catch to this analogy: in the Legion programming
model two different logical regions are not necessarily independent from each other.  In
order to do the dependence detection we rely on the region tree data structure to be
able to detect dependences between tasks which require different logical regions.

@@

Once we have found the root region for the requested region, we then proceed to walk
down the region tree from the node corresponding to the root region to the node
for the requested region.  At each node, we determine if any potential dependences exist.

For region nodes we first check to see if the required region has any dependences on
tasks currently using the logical region by performing the check on coherence properties
described in section \ref{subsec:cohdep}.  If a conflict is discovered
we record that the task must wait on the previous task before executing.  We also check
to see if any of the partitions of the region have been marked open.  If they have
we must {\em close} the partitions to detect any possible conflicts

@@

\subsubsection{Instance Management}
\label{subsec:instmang}
